A survey of HTTP caching implementations on the open Semantic Web

### 2015-01-20 sanity/paper1.tex
Jonas proofreading
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1023 43ef1190-90fd-4dd7-a769-d0121f02b80e

\documentclass{llncs}
%\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{verbatim}
\usepackage{graphicx}
%\usepackage{wrapfig}
\usepackage{appendix}

@@

RFC7234 defines two important headers: \httph{Expires}, whose value is
a date and time of when the response is considered stale, and
therefore should not be used; and \httph{Cache-Control}, which allows
detailed control of the cache, including a \httph{max-age} field,
which gives the time in seconds for how long the the response may be
used from the time of the request. \httph{max-age} takes precedence
over \httph{Expires}. In the following \emph{freshness lifetime} is
understood as the number of seconds that the response may be used
without contacting the origin server. Ideally, the calculation of the
freshness lifetime should be based of the above, we therefore shall
refer to this as ``standards-compliant caching''. However it can also
be based on heuristics, and RFC7234, Section~4.2.2 provides some loose
constraints for such practice, as well as a suggestion for a useful
heuristic. This heuristic is based on a fraction of the time lapsed
between the current time and the modification time given in the
\httph{Last-Modified} header. This approach therefore still requires
the Web server to be cooperative to be successful, but commonly, Web
servers can track this, for example if RDF is served from a file
system, the file modification time is used.

@@

While these studies have little overlap with the present paper, they
underline the importance of understanding the current deployment and
future potential. In some of the related work, it is shown that
caching does not necessarily give tangible benefits. Yet, we shall
assume that sharing the meta data required for caching outside of the
server is desirable, and that it is possible in most cases. We shall
see that it most likely will be beneficial in cases that do not
benefit from caching today.

@@

We make several approaches to ensure that we visit a large and
representative section of the open Semantic Web. We take SPARQL
Endpoints from the SPARQLES survey \cite{buil2013sparql}, vocabularies
from Linked Open Vocabularies (LOV) \cite{lov2} and prefix.cc, and
we augment these data with spidered data from the Billion Triple
Challenge 2014 \cite{btc-2014} dataset.

@@

First, we investigated whether certain server implementations provided
better support for caching than others. To do this, we formulated
SPARQL queries to examine the \httph{Server} headers of successful
responses, with optional clauses matching the standard-compliant
computed freshness lifetime (which is the ideal) as well as whether
the response had other promising features that may assist caching such
as modification time, certain predicates, etc.%\todo{promising}

@@

For a more systematic approach, we wish to test the hypothesis that
some servers are better configured to support caching than others.
Using the methodology given in Appendix~\ref{app:stats}, we find in
both the cases of standard-compliant freshness lifetime and for the other
promising features, the test reports \pvalue = 0.0001. We can conclude
that it is highly likely that some servers are better at exposing
cache headers than others. Unfortunately, since most \httph{Server}
headers only contain generic values, little can be learnt about these
implementations. We note, however, that DBPedia exposes
standards-compliant freshness lifetime of 604800 seconds (i.e. 1 week) for
both LOD and SPARQL endpoints.

@@

Of the 2965 resources, 405 returned valid headers, but 114 did so to
prohibit caching of the response, and 3 contained conflicting
headers, i.e. set an freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occurred, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching. 
Note that the freshness lifetime is 0 whenever caching is prohibited.

@@

\begin{figure}[ht]
  \centerline{%
    \includegraphics[width=.9\textwidth]{hardall.pdf}}
  \caption{Barplot counting all standard-compliant freshness lifetimes
    found, with the percentage of occurrences indicated on the
    bars. The first bin are the cases where caching is explicitly
    prohibited. The next bins are for lifetimes, where the values are
    grouped if they are on the order of seconds, minutes, hours, etc,
    i.e. the second bin counts the lifetimes in the interval [1,59]
    seconds, etc.}
  \label{fig:hardall}
\end{figure}

@@

We also note that prohibition of SPARQL results caching is rare,
amongst the servers that expose caching headers, it is common
that the result may be cached for some minutes.

@@

We next consider the simple heuristic freshness lifetime as suggested
RFC7234, Section~4.2.2 and mentioned in the introduction.

@@

We found that errors are commonplace, but that they does not usually
pertain to the caching headers. We found a small number of
self-contradictory \httph{Cache-Control} headers, and some servers
that set conditional request response headers, that could not support
conditional requests.

@@

In spite of this, most of the Semantic Web is unhelpful in spite of
being slowly changing.

@@

In this work, we have only explored the cases where the server is
cooperative, in the sense that message data or metadata provides at
least a hint of a resource's cacheability. We also noted that the
majority of the Semantic Web is unhelpful. Therefore, an interesting
direction is to learn the change frequency of resources to use in a
heuristic freshness lifetime estimation. However, such work should
operate within the loose constraints of RFC7234, Section~4.2.2, and
should find its niche when the other techniques described in this paper
are unavailable. Once this is done, the correctness of caching headers
should be assessed, possibly using contigency tables similar to those
in Appendix~\ref{app:stats}.

@@

In many cases, setting a reasonable cache headers is straightforward,
and should be done by data owners. Framework authors should make it
easy to set the expected lifetime, heeding the metadata association
good practice recommendation of \cite{Jacobs:04:AWW}. If the author is
unable to influence HTTP headers, they should set a
\rdfterm{dct:valid} time into the future and make use of
\rdfterm{dct:modified}.

@@

\paragraph*{Acknowledgements} The author would like to thank Jürgen
Umbrich and Martin Giese for careful review and critical comments, and
Gregory Todd Williams for promptly solving issues in the underlying
libraries used in this study, and Axel Polleres for encouraging comments.

@@

We found in the data a large number of ontologies that consist of many
information resources with just a few triples in each. Since they
appear to be produced by the same software, usually Semantic
Mediawiki, we assume that they are configured with a single setup, and
thus we will merely sample these resources.

We continued to also sample the HTTP headers gathered in the BTC2014
dataset. 
%as they had recorded some of the relevant headers,
%\httph{Expires}, \httph{Last-Modified} and \httph{ETag}. 
First, we traversed the files with a simple UNIX \texttt{grep} to find
the resources that had reported one of the RDF serializations as
content type. We then traversed this list, first discarded the
resources that did not have a valid IRI (this amounted to just 3273
resources). For the resources we found, as well as for the resources
that was of \rdfterm{rdf:type owl:Ontology} above, we kept one
resource per hostname, with the exception of the popular blogging
platforms Livejournal and SAPO, where each blog has their own host and
they expose FOAF data. For those, we only kept one hostname.

@@

Then, the first request would be made, and a selection of the
resulting HTTP headers recorded in an per-host NQuads file. For this
purpose, we developed and released a module
RDF::Generator::HTTP\footnote{https://metacpan.org/release/KJETILK/RDF-Generator-HTTP-0.003}
to CPAN. We recorded whether the conditional request showed that the
BTC data were still fresh, and if it was, we retrieved the current
data, as we had not coupled the headers to the body in our original
retrieval.  Based on the resulting headers, we let
libwww\footnote{https://metacpan.org/release/GAAS/libwww-perl-6.04}
calculate both standards-compliant and simple heuristic freshness
lifetime.

@@

The hypothesis tests in this paper is implemented using \emph{contigency
  tables} (see e.g. \cite{kn:bj}). This formalism is suited to see if the
distribution of  \httph{Server} headers are different for those implementations that
offer caching headers from those that don't. Intuitively, we expect
these distributions to be similar, ``long-tail'' distributions, i.e. a
handful of servers are used by a large number of projects, and then it
falls off rapidly, and so, some servers are used only by very
few. Likewise, it is to be expected that only a few projects have
given caching enough attention, but that they account for the majority
of the support. The question is if the presence of caching headers is
merely a matter of proportion, or if there are some that have given it
more attention, but still is in relatively little use.

>>>>

\documentclass{llncs}
\usepackage[english,british]{babel}
%\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{appendix}

@@

RFC7234 defines two important headers: \httph{Expires}, whose value is
a date and time of when the response is considered stale, and
therefore should not be used; and \httph{Cache-Control}, which allows
detailed control of the cache, including a \httph{max-age} field,
which gives the time in seconds for how long the the response may be
used from the time of the request. \httph{max-age} takes precedence
over \httph{Expires}. In the following \emph{freshness lifetime} is
understood as the number of seconds that the response may be used
without contacting the origin server. Ideally, the calculation of the
freshness lifetime should be based of the above, we therefore shall
refer to this as ``standards-compliant caching''. However it can also
be based on heuristics, and RFC7234, Section~4.2.2 provides some loose
constraints for such practice, as well as a suggestion for a useful
heuristic. This heuristic is based on a fraction of the time lapsed
between the current time and the modification time given in the
\httph{Last-Modified} header. This approach therefore still requires
the Web server to be cooperative to be successful. Commonly, Web
servers can track this, for example if RDF is served from a file
system the file modification time is used.

@@

While these studies have little overlap with the present paper, they
underline the importance of understanding the current deployment and
future potential. In some of the related work, it is shown that
caching does not necessarily give tangible benefits. Yet, we shall
assume that sharing the metadata required for caching outside of the
server is desirable, and that it is possible in most cases. We shall
see that it most likely will be beneficial in cases that do not
benefit from caching today.

@@

We make several approaches to ensure that we visit a large and
representative section of the open Semantic Web. We take SPARQL
Endpoints from the SPARQLES survey \cite{buil2013sparql}, vocabularies
from Linked Open Vocabularies (LOV) \cite{lov2} and prefix.cc, and
we augment these data with spidered data from the Billion Triple
Challenge (BTC) 2014 \cite{btc-2014} dataset.

@@

First, we investigated whether certain server implementations provided
better support for caching than others. To do this, we formulated
SPARQL queries to examine the \httph{Server} headers of successful
responses, with optional clauses matching the standards-compliant
computed freshness lifetime (which is the ideal) as well as whether
the response had other promising features that may assist caching such
as modification time, certain predicates, etc.%\todo{promising}

@@

For a more systematic approach, we wish to test the hypothesis that
some servers are better configured to support caching than others.
Using the methodology given in Appendix~\ref{app:stats}, we find in
both the cases of standards-compliant freshness lifetime and for the other
promising features, the test reports \pvalue = 0.0001. We can conclude
that it is highly likely that some servers are better at exposing
cache headers than others. Unfortunately, since most \httph{Server}
headers only contain generic values, little can be learnt about these
implementations. We note, however, that DBPedia exposes
standards-compliant freshness lifetime of 604800 seconds (i.e. 1 week) for
both LOD and SPARQL endpoints.

@@

Of the 2965 resources, 405 returned valid headers, but 114 did so to
prohibit caching of the response, and 3 contained conflicting
headers, i.e. set a freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occurred, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching. 
Note that the freshness lifetime is 0 whenever caching is prohibited.

@@

\begin{figure}[ht]
  \centerline{%
    \includegraphics[width=.9\textwidth]{hardall.pdf}}
  \caption{Barplot counting all standards-compliant freshness lifetimes
    found, with the percentage of occurrences indicated on the
    bars. The first bin are the cases where caching is explicitly
    prohibited. The next bins are for lifetimes, where the values are
    grouped if they are on the order of seconds, minutes, hours, etc,
    i.e. the second bin counts the lifetimes in the interval [1,59]
    seconds, etc.}
  \label{fig:hardall}
\end{figure}

@@

We also note that prohibition of SPARQL results caching is rare.
Amongst the servers that expose caching headers it is common
that the result may be cached for some minutes.

@@

We next consider the simple heuristic freshness lifetime as suggested in
RFC7234, Section~4.2.2 and mentioned in the introduction.

@@

We found that errors are commonplace, but that they do not usually
pertain to the caching headers. We found a small number of
self-contradictory \httph{Cache-Control} headers, and some servers
that set conditional request response headers, but could not support
conditional requests.

@@

In spite of this, most of the Semantic Web is unhelpful in even
though it is changing slowly.

@@

In this work, we have only explored the cases where the server is
cooperative, in the sense that message data or metadata provides at
least a hint of a resource's cacheability. We also noted that the
majority of the Semantic Web is unhelpful. Therefore, an interesting
direction is to learn the change frequency of resources to use in a
heuristic freshness lifetime estimation. However, such work should
operate within the loose constraints of RFC7234, Section~4.2.2, and
should find its niche when the other techniques described in this paper
are unavailable. Once this is done, the correctness of caching headers
should be assessed, possibly using contingency tables similar to those
in Appendix~\ref{app:stats}.

@@

In many cases, setting reasonable cache headers is straightforward,
and should be done by data owners. Framework authors should make it
easy to set the expected lifetime, heeding the metadata association
good practice recommendation of \cite{Jacobs:04:AWW}. If the author is
unable to influence HTTP headers, they should set a
\rdfterm{dct:valid} time into the future and make use of
\rdfterm{dct:modified}.

@@

\paragraph*{Acknowledgements} The author would like to thank Jürgen
Umbrich and Martin Giese for careful review and critical comments, and
Gregory Todd Williams for promptly solving issues in the underlying
libraries used in this study, Axel Polleres for encouraging comments
and Jonas Smedegaard for proofreading.

@@

We found in the data a large number of ontologies that consist of many
information resources with just a few triples in each. Since they
appear to be produced by the same software, usually Semantic
Mediawiki, we assume that they are configured with a single setup, and
thus we merely sampled these resources.

We continued to also sample the HTTP headers gathered in the BTC2014
dataset. 
%as they had recorded some of the relevant headers,
%\httph{Expires}, \httph{Last-Modified} and \httph{ETag}. 
First, we traversed the files with a simple UNIX \texttt{grep} to find
the resources that had reported one of the RDF serializations as
content type. We then traversed this list, first discarded the
resources that did not have a valid IRI (this amounted to just 3273
resources). For the resources we found, as well as for the resources
that was of \rdfterm{rdf:type owl:Ontology} above, we kept one
resource per hostname, with the exception of the popular blogging
platforms Livejournal and SAPO, where each blog has their own host and
they expose FOAF data. For those, we only kept one hostname, since
they are likely configured similarly.

@@

Then, the first request would be made, and a selection of the
resulting HTTP headers recorded in an per-host NQuads file. For this
purpose, we developed and released a module
RDF::Generator::HTTP\footnote{https://metacpan.org/release/KJETILK/RDF-Generator-HTTP-0.003}
to CPAN. We recorded whether the conditional request showed that the
BTC2014 data were still fresh, and if it was, we retrieved the current
data, as we had not coupled the headers to the body in our original
retrieval.  Based on the resulting headers, we let
libwww\footnote{https://metacpan.org/release/GAAS/libwww-perl-6.04}
calculate both standards-compliant and simple heuristic freshness
lifetime.

@@

The hypothesis tests in this paper is implemented using \emph{contingency
  tables} (see e.g. \cite{kn:bj}). This formalism is suited to see if the
distribution of  \httph{Server} headers are different for those implementations that
offer caching headers from those that don't. Intuitively, we expect
these distributions to be similar, ``long-tail'' distributions, i.e. a
handful of servers are used by a large number of projects, and then it
falls off rapidly, and so, some servers are used only by very
few. Likewise, it is to be expected that only a few projects have
given caching enough attention, but that they account for the majority
of the support. The question is if the presence of caching headers is
merely a matter of proportion, or if there are some that have given it
more attention, but still is in relatively little use.
